---
title: "Analyse Datasets and Models"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

rm(list = ls())

library(tidyverse)
library(readr)
library(tibble)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(caret)
library(data.table)
library(matrixStats)
library(tictoc)                 # Used for timing of code
library(MazamaCoreUtils)        # Used for logging messages to a file

source("./Code/utilities.R")
source("./Code/Train/read_dataset.R")
source("./Code/Train/process_config.R")


```


```{r, initialise}

READ_FRESH = TRUE

proj_path   <- file.path(".")
data_path   <- file.path(proj_path, "DataSets")
log_path    <- file.path(proj_path, "Logs")
rdata_path  <- file.path(proj_path, "rData")
config_path <- file.path(proj_path, "config")


data_sets <- list(
  "Baron_Mouse",  # 1,886 cells, 14,861 genes
  "Baron_Human",  # 8,569 cells, 17,499 genes
  "Muraro",       # 2,122 cells, 18,915 genes
  "SegerStolpe",  # 2,133 cells, 22,757 genes
  "Xin",          # 1,449 cells, 33,889 genes
  "Zhou",         # 2,554 cells, 19,046 genes
  "Bo",           # 516   cells, 56,924 genes
  "Petropoulos1"  # 1,529 cells, 26,178 genes 
)

gene_names  <- list()
class_names <- list()
class_freq  <- list()
top_genes   <- list()

ds_table    <- data.frame()

```


# Generate data overview (cells, genes, classe - before and after small cell correction) 

```{r}

for (dataset_name in data_sets) {
  
  cat(sprintf("\n\n\nProcessing %s\n\n", dataset_name))

  # Set the data and model name and read the data
  ret                <- read_dataset(data_path, dataset_name)
  data               <- ret$data
  labels             <- ret$labels
  
  dim_uncorrected <- dim(data)
  class_uncorrected  <- table(labels$ident)
  
  print(dim_uncorrected)
  print(class_uncorrected)
  
  ########################################
  # Overrule settings from config file
  ########################################

  min_class_size      <- 10 
  minCount            <- 0 
  maxCount            <- -1
  minFeature          <- 0 
  maxFeature          <- -1   

  ret              <- process_config(min_class_size, minCount, maxCount, minFeature, maxFeature, data, labels)
  data             <- ret$data
  labels           <- ret$labels
  
  dim_corrected    <- dim(data)
  class_corrected  <- table(labels$ident)

  print(dim_corrected)
  print(class_corrected)
  
  ds_row <- data.frame(Cells       = dim_uncorrected[1],
                       Genes       = dim_uncorrected[2],
                       Classes     = length (class_uncorrected),
                       Cells_Cor   = dim_corrected[1],
                       Genes_Cor   = dim_corrected[2],
                       Classes_Cor = length (class_corrected))
  rownames(ds_row)[1] <- datset_name
  ds_table            <- rbind(ds_table, ds_row)
  print(ds_table)
}

ds_table

```

```{r}
dataset_name <- "AMB"
cat(sprintf("\n\n\nProcessing %s\n\n", dataset_name))

# Set the data and model name and read the data
ret                <- read_dataset(data_path, dataset_name)
data               <- ret$data
labels             <- ret$labels

dim_uncorrected             <- dim(data)
class_ident_uncorrected     <- table(labels$ident)
class_subclass_uncorrected  <- table(labels$Subclass)
class_cluster_uncorrected   <- table(labels$cluster)

print(dim_uncorrected)
print(class_ident_uncorrected)
print(class_subclass_uncorrected)
print(class_cluster_uncorrected)

saved_data   <- data
saved_labels <- labels

########################################
# Overrule settings from config file
########################################

min_class_size      <- 10 
minCount            <- 0 
maxCount            <- -1
minFeature          <- 0 
maxFeature          <- -1   


# Process the ident level
data               <- saved_data 
labels             <- saved_labels 
labels_to_process  <- select(labels, "ident", "cell_id")
ret                <- process_config(min_class_size, minCount, maxCount, minFeature, maxFeature, data, labels_to_process)
data               <- ret$data
labels             <- ret$labels

dim_corrected          <- dim(data)
class_ident_corrected  <- table(labels$ident)

print(dim_corrected)
print(class_ident_corrected)

ds_row <- data.frame(Cells       = dim_uncorrected[1],
                     Genes       = dim_uncorrected[2],
                     Classes     = length (class_ident_uncorrected),
                     Cells_Cor   = dim_corrected[1],
                     Genes_Cor   = dim_corrected[2],
                     Classes_Cor = length (class_ident_corrected))
rownames(ds_row)[1] <- "AMB 1"
ds_table            <- rbind(ds_table, ds_row)
print(ds_table)
  

# Process the subclass level
data               <- saved_data 
labels             <- saved_labels
labels_to_process  <- select(labels, "Subclass", "cell_id")
colnames(labels_to_process)[1]  <-  "ident"
ret                <- process_config(min_class_size, minCount, maxCount, minFeature, maxFeature, data, labels_to_process)
data               <- ret$data
labels             <- ret$labels

dim_subclass_corrected          <- dim(data)
class_subclass_corrected        <- table(labels$ident)

print(dim_corrected)
print(class_subclass_corrected)

ds_row <- data.frame(Cells       = dim_uncorrected[1],
                     Genes       = dim_uncorrected[2],
                     Classes     = length (class_subclass_uncorrected),
                     Cells_Cor   = dim_subclass_corrected[1],
                     Genes_Cor   = dim_subclass_corrected[2],
                     Classes_Cor = length (class_subclass_corrected))

rownames(ds_row)[1] <- "AMB 2"
ds_table            <- rbind(ds_table, ds_row)

print(ds_table)

# Process the cluster level
data               <- saved_data 
labels             <- saved_labels
labels_to_process  <- select(labels, "cluster", "cell_id")
colnames(labels_to_process)[1]  <-  "ident"
ret                <- process_config(min_class_size, minCount, maxCount, minFeature, maxFeature, data, labels_to_process)
data               <- ret$data
labels             <- ret$labels

dim_cluster_corrected          <- dim(data)
class_cluster_corrected        <- table(labels$ident)

print(dim_corrected)
print(class_cluster_corrected)

ds_row <- data.frame(Cells       = dim_uncorrected[1],
                     Genes       = dim_uncorrected[2],
                     Classes     = length (class_cluster_uncorrected),
                     Cells_Cor   = dim_cluster_corrected[1],
                     Genes_Cor   = dim_cluster_corrected[2],
                     Classes_Cor = length (class_cluster_corrected))

rownames(ds_row)[1] <- "AMB 3"
ds_table            <- rbind(ds_table, ds_row)

print(ds_table)
```





It is not necessary to read in the data again, but just to avoidinconsistency. The above was done becaise we needed bith umporcesses and processed. For the remainder in the file you need only processed.  


```{r, read data}

if (READ_FRESH) {

  # Read the data
  
  for (ii in 1:length(data_sets)) {
    dataset_name <- data_sets[[ii]][1]
    cat(sprintf("%s\n", dataset_name))
    
    ret    <- load_single_data(rdata_path, dataset_name)
    data   <- ret$data
    labels <- ret$labels

    gene_names[ii]  <- list(sort(colnames(data)))
    class_names[ii] <- list(sort(str_to_lower(unique(labels[,1]))))
    class_freq[ii]  <- list(table(labels$ident))
    
    # To read in the top_genes, you can use any of the models for that dataset, svnLinear and pca is a good combination, but any will do
    ret <- load_single_model(rdata_path, dataset_name, "svmLinear", "pca")
    top_genes[ii]      <- list(ret$model_information[["top_genes"]])
  }
  
  names(class_names) <- data_sets
  names(class_freq)  <- data_sets
  names(gene_names)  <- data_sets
 
  # Save the genes per dataset
  save(gene_names, class_names, class_freq, top_genes, 
       file=file.path(rdata_path, "dataset_info.rData"))
  
} else {
  load(file.path(rdata_path, "dataset_info.rData"))
}

```


# Report on datasets

```{r}

for (i in 1:length(gene_names)) {
  cat(sprintf("Corrected length of %20s is %d\n", names(gene_names[i]), length(gene_names[[i]])))
}

```


```{r, class overview}

# Get a list of all class names
all_class = class_names[[1]]
for (c in class_names[-1]) {
  all_class = union(all_class, unlist(c))
}
nr_classes <- length(all_class)

# Create an empty dataframe with as rows the classes and as columns the data zsets
new_col        <- rep(0, nr_classes)
class_overview <- data.frame(start = new_col)

for (c in data_sets) {
  class_overview[c] <- new_col 
}
row.names(class_overview) = all_class
class_overview <- class_overview[,-1]

# Fill the data frame

for (i in 1:length(class_freq)) {
  col_name <- names(class_freq[i])
  for (j in 1:length(class_freq[[i]])) {
    row_name = str_to_lower(names(class_freq[[i]][j]))
    class_overview[row_name, col_name]  = class_freq[[i]][j]
  }
}
class_overview <- class_overview[order(row.names(class_overview)),] 
head(class_overview, 30)

class_info <- data.frame(
  nCell   = colSums(class_overview),
  nClass  = (dim(class_overview)[1] - colCounts(as.matrix(class_overview), value = 0))
)

class_info

```

# How many genes do the data sets have in common?

```{r, common genes}

# How many genes do the data sets have in common?
shared_genes <- gene_names[[1]]
for (i in 2:length(gene_names)) {
  shared_genes = intersect(shared_genes, unlist(gene_names[[i]]))
}
nr_shared_genes <- length(shared_genes)
nr_shared_genes

```


# How many genes are there in total?

```{r, total genes}

total_genes <- character()
cat(sprintf("Length of set is %d\n", length(total_genes)))
for (i in 1:length(gene_names)) {
  total_genes = union(total_genes, unlist(unlist(gene_names[[i]])))
  cat(sprintf("After adding %s Length of set is %d\n", names(gene_names[i]), length(total_genes)))
}
nr_total_genes <- length(total_genes)
nr_total_genes

```


# Do the top genes appear in all data sets?

```{r, intersect top genes - pancreatic}

size_of_top_genes <- 1000

cat(sprintf("How many of the %d top genes occur in other datasets? \n\n", size_of_top_genes))

for (i in 1:5) { 
  for (j in 1:5) {
    overlap <- length(intersect(top_genes[[i]][1:size_of_top_genes],
                                gene_names[[j]]))
    cat(sprintf("Of the top %d genes of %-12s %4d occur in the full %-12s dataset\n", size_of_top_genes, data_sets[i], overlap, data_sets[j]))
  }
  cat(sprintf("\n"))
}

```
```{r, intersect top genes - embryonic}

size_of_top_genes <- 1000

cat(sprintf("How many of the %d top genes occur in other datasets? \n\n", size_of_top_genes))

for (i in 6:8) { 
  for (j in 6:8) {
    overlap <- length(intersect(top_genes[[i]][1:size_of_top_genes],
                                gene_names[[j]]))
    cat(sprintf("Of the top %d genes of %-12s %4d occur in the full %-12s dataset\n", size_of_top_genes, data_sets[i], overlap, data_sets[j]))
  }
  cat(sprintf("\n"))
}

```



# Are the top_genes very different in the various datasets?

```{r, topgenes in datasets}
# size_of_top_genes <- 1000
# 
# cat(sprintf("Analysis of occurance of the top %d genes used in the model in the various data sets.\n\n", size_of_top_genes))
# 
# for (i in 1:length(data_sets)) {
#   for (j in 1:length(data_sets)) {
#     overlap <- length(intersect(top_genes[[i]][1:size_of_top_genes],
#                                 top_genes[[j]][1:size_of_top_genes]))
#     cat(sprintf("The overlap of the %d top genes in model %-12s with the top genes in dataset %-12s is %d\n", 
#                 size_of_top_genes, data_sets[i], data_sets[j], overlap))
#   }
#   cat("\n")
# }
```


```{r, topgenes in datasets - pancreatic}
cat(sprintf("Analysis of occurance of the top %d genes used in the model in the pancreas data sets.\n\n", size_of_top_genes))

for (i in 1:5) { 
  for (j in 1:5) {
    overlap <- length(intersect(top_genes[[i]][1:size_of_top_genes],
                                top_genes[[j]][1:size_of_top_genes]))
    cat(sprintf("The overlap of the %d top genes in model %-12s with the top genes in dataset %-12s is %d\n", 
                size_of_top_genes, data_sets[i], data_sets[j], overlap))
  }
  cat("\n")
}
```


```{r, topgenes in datasets - embryonic}
cat(sprintf("Analysis of occurance of the top %d genes used in the model in the embryonic data sets.\n\n", size_of_top_genes))

for (i in 6:8) { 
  for (j in 6:8) {
    overlap <- length(intersect(top_genes[[i]][1:size_of_top_genes],
                                top_genes[[j]][1:size_of_top_genes]))
    cat(sprintf("The overlap of the %d top genes in model %-12s with the top genes in dataset %-12s is %d\n", 
                size_of_top_genes, data_sets[i], data_sets[j], overlap))
  }
  cat("\n")
}
```
