---
title: "Review AMB data set"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include = FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
rm(list = ls())
set.seed(1234)
```


# Set key use parameters

```{r}

# Set datasetname

dataset_name     <- "AMB"

# Select preprocessing

preproc_method   <- "pca"
preproc_method   <- "nopca"
preproc_method   <- "logpca"

preproc_method   <- "logpca"
```

```{r, libraries, message=FALSE}

library(readr)
library(tibble)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(Seurat)
library(caret)
library(data.table)
library(matrixStats)
library(tictoc)
library(MazamaCoreUtils)        # Used for logging messages to a file
library(ggraph)
library(data.tree)
library(DiagrammeR)

```


```{r, initialise, message=FALSE}

PCA_Threshold   <- 0.90         # The percentage variance explained by the PCA
                                # A high percentage requires more PC dimensions to be considered

features_limit  <- 1000         # The number of features that will be considered duringthe PCA
                                # A high number use sthe richness of the data set optimally, but makes processing slo

VERBOSE         <- TRUE         # If TRUE generates analysis data
SEURAT_VERBOSE  <- FALSE
LOGGER          <- FALSE

# Summary results will be kept here 
report_out <- data.frame(Method     = character(),
                         ModelData  = character(),
                         TestData   = character(),
                         Accuracy   = numeric(),
                         Confidence = numeric())

# Information of all the runs will be gathered in the class_summaries list  
class_summaries <- list()

proj_path        <- file.path(".")
data_path        <- file.path(proj_path, "DataSets")
log_path         <- file.path(proj_path, "Logs")
rdata_path       <- file.path(proj_path, "rData")
config_path      <- file.path(proj_path, "config")

logger.setup (infoLog = file.path(log_path, "amb.log"))

```


```{r, set preprocess options}

if (preproc_method == "pca")  {
  log_preprocess   <- FALSE
  caret_preprocess_method <- c('scale', 'center', 'pca')
} else if (preproc_method == "nopca")  {
  log_preprocess   <- FALSE
  caret_preprocess_method <- c('scale', 'center')
} else if (preproc_method == "logpca")  {
  log_preprocess   <- TRUE
  caret_preprocess_method <- c('scale', 'center', 'pca')
}

```


# Read in the data

```{r, Read}

# Start timer
logger.info("Info: Started reading data from dataset %s", dataset_name)
tic(msg = "", quiet = TRUE)

dataset_filename <- paste0(dataset_name, "_model_compare")

READ_FRESH <- FALSE

if (READ_FRESH) {
  # Read the data and labels and format names
  data      <- read.csv(file.path(data_path, dataset_name, "Data.csv"), 
                        stringsAsFactors = FALSE, 
                        row.names = 1)
  labels    <- read.csv(file.path(data_path, dataset_name, "Labels.csv"), 
                        stringsAsFactors = FALSE)
  
  save(data, labels, file = file.path(rdata_path, "amb_data.rData"))
} else {
  load(file.path(rdata_path, "amb_data.rData"))
}

# Log transform if requested
# Make sure there are no blanks or dashes or slashes in class names
labels      <- apply(labels, 2, function(x) str_replace_all( x, "[ -/]", "."))
labels      <- as.data.frame(labels)
save_data   <- data
save_labels <- labels

# Report timer
toc(log=TRUE, quiet = TRUE)
logger.info("Info: Finished reading data from dataset %s%s", dataset_name, tic.log()[[1]])
tic.clearlog()

```

# Inspect structure

## Original

```{r}

labels %>%
  group_by( cluster) %>%
  summarise(n=n())  %>% 
filter( n > 10) 
  
```

## Class level

```{r}

l1 <- labels %>%
  group_by(Class, Subclass) %>%
  summarise(NumberOfSubclass=n())  %>%
  group_by(Class) %>%
  summarise(NumberOfSubclass=n()) 

l2 <- labels %>%
  group_by(Class) %>%
  summarise(CellsInClass = n()) 

l1 <- column_to_rownames(l1, var = "Class")
l2 <- column_to_rownames(l2, var = "Class")
cbind(l1,l2)

```



## Subclass level GABAergic

```{r}

l1 <- labels %>%
  filter(Class == "GABAergic") %>%
  group_by(Class, Subclass, cluster) %>%
  summarise(n = n()) %>%
  #filter( n > 10) %>%
  group_by(Subclass) %>%
  summarise(NumberOfCluster = n()) %>%
  arrange (Subclass) 

l2 <- labels %>%
  filter(Class == "GABAergic") %>%
  group_by(Subclass) %>%
  summarise(CellsInCluster = n()) %>%
  #ilter( CellsInCluster > 10) %>%
  arrange (Subclass)

l1 <- column_to_rownames(l1, var = "Subclass")
l2 <- column_to_rownames(l2, var = "Subclass")
l  <- cbind(l1,l2)
l

```


## Subclass level Glutamatergic

```{r}

l1 <- labels %>%
  filter(Class == "Glutamatergic") %>%
  group_by(Class, Subclass, cluster) %>%
  summarise(n = n()) %>%
  #filter( n > 10) %>%
  group_by(Subclass) %>%
  summarise(NumberOfCluster = n()) %>%
  arrange (Subclass) 

l2 <- labels %>%
  filter(Class == "Glutamatergic") %>%
  group_by(Subclass) %>%
  summarise(CellsInCluster = n()) %>%
  #filter( CellsInCluster > 10) %>%
  arrange (Subclass)

l1 <- column_to_rownames(l1, var = "Subclass")
l2 <- column_to_rownames(l2, var = "Subclass")
l  <- cbind(l1,l2)
l

```

## Subclass level Endothelial

```{r}

l1 <- labels %>%
  filter(Class == "Endothelial") %>%
  group_by(Class, Subclass, cluster) %>%
  summarise(n = n()) %>%
  #filter( n > 10) %>%
  group_by(Subclass) %>%
  summarise(NumberOfCluster = n()) %>%
  arrange (Subclass) 

l2 <- labels %>%
  filter(Class == "Endothelial") %>%
  group_by(Subclass) %>%
  summarise(CellsInCluster = n()) %>%
  #filter( CellsInCluster > 10) %>%
  arrange (Subclass)

l1 <- column_to_rownames(l1, var = "Subclass")
l2 <- column_to_rownames(l2, var = "Subclass")
l  <- cbind(l1,l2)
l

```

## Subclass level Non.Neuronal

```{r}

l1 <- labels %>%
  filter(Class == "Non.Neuronal") %>%
  group_by(Class, Subclass, cluster) %>%
  summarise(n = n()) %>%
  #filter( n > 10) %>%
  group_by(Subclass) %>%
  summarise(NumberOfCluster = n()) %>%
  arrange (Subclass) 

l2 <- labels %>%
  filter(Class == "Non.Neuronal") %>%
  group_by(Subclass) %>%
  summarise(CellsInCluster = n()) %>%
  #filter( CellsInCluster > 10) %>%
  arrange (Subclass)

l1 <- column_to_rownames(l1, var = "Subclass")
l2 <- column_to_rownames(l2, var = "Subclass")
l  <- cbind(l1,l2)
l

```


```{r}

labels %>%
  filter(Class == "Glutamatergic") %>%
  group_by(Subclass) %>%
  summarise(CellsInCluster = n()) %>%
  filter( CellsInCluster > 10) %>%
  arrange (Subclass)

```


```{r}

labels %>%
  group_by(cluster) %>%
  summarise(n = n()) %>%
  arrange (cluster)

```


```{r}

labels %>%
  group_by(Class, Subclass, cluster) %>%
  summarise(n = n())  %>%
  filter( Class !=  "Glutamatergic" & Class !=  "GABAergic") %>%
  arrange (Class, Subclass, cluster)

```


```{r}

l <- labels
l$amb        <- rep("amb", dim(l)[1])
l            <- l[, c("amb", "Class")]
l$pathString <- do.call(paste,c(l,sep="/"))
l.graph      <- data.tree::as.Node(l)
plot(l.graph)

```


```{r}

l <- labels
l$amb        <- rep("amb", dim(l)[1])
l            <- l[, c("amb", "Class", "Subclass")]
l$pathString <- do.call(paste,c(l,sep="/"))
l.graph      <- data.tree::as.Node(l)
plot(l.graph)

```


```{r}

l <- labels
l <- l %>%
  filter(Class == "Glutamatergic") %>%
  select(Class, Subclass)
l$pathString <- do.call(paste,c(l,sep="/"))
l.graph      <- data.tree::as.Node(l)
plot(l.graph)

```


```{r}

l <- labels
l <- l %>%
  filter(Class == "GABAergic") %>%
  select(Class, Subclass)
l$pathString <- do.call(paste,c(l,sep="/"))
l.graph      <- data.tree::as.Node(l)
plot(l.graph)

```


```{r}

l <- labels
l <- l %>%
  filter(Subclass == "Vip") %>%
  select(Subclass, cluster)
l$pathString <- do.call(paste,c(l,sep="/"))
l.graph      <- data.tree::as.Node(l)
plot(l.graph)

```

# Find the Variable Genes 

```{r, variable_features, fig.height=5, message = FALSE, echo=FALSE}

# Start timer
logger.info("Info: Started determining variable genes" )
tic(msg = "", quiet = TRUE)

seurat_object <- CreateSeuratObject(counts    = t(data),
                                    meta.data = as.data.frame(labels),
                                    min.cells = 5)

seurat_object <- NormalizeData(seurat_object, 
                               normalization.method = "LogNormalize", 
                               scale.factor = 1000000)

seurat_object <- FindVariableFeatures(object           = seurat_object, 
                                      selection.method = 'vst',
                                      verbose          = TRUE,
                                      nfeatures        = dim(data)[2])

# Identify the most highly variable genes
top_genes      <- VariableFeatures(seurat_object)

# Show composition
cat(sprintf("Information on dataset after variable gene determination of dataset %s\n", dataset_name))
cat(sprintf("Removed %d genes that apparently are not very variable \n\n", dim(data)[2] - length(top_genes) )) 

data = data[ , top_genes]
cat(sprintf("\nThere are %d cells and %d features.\n", dim(data)[1], dim(data)[2]))

# Report timer
toc(log = TRUE, quiet = TRUE)
logger.info("Info: Ended determining variable genes%s", tic.log()[[1]])
tic.clearlog()

```


```{r, Class information, fig.height=5, message = FALSE, echo=FALSE}

# Start timer
logger.info("Info: Started classinfo")
tic(msg = "", quiet = TRUE)

cat(sprintf("\nClass information for 'Class'\n"))
for (i in 1:length(table(labels$Class))) {
  cat(sprintf("%-25s %d\n", names(table(labels$Class)[i]), table(labels$Class)[i]))
}
cat(sprintf("\n"))

cat(sprintf("\nClass information for 'Subclass'\n"))
for (i in 1:length(table(labels$Subclass))) {
  cat(sprintf("%-25s %d\n", names(table(labels$Subclass)[i]), table(labels$Subclass)[i]))
}
cat(sprintf("\n"))

cat(sprintf("\nClass information for 'cluster'\n"))
for (i in 1:length(table(labels$cluster))) {
  cat(sprintf("%-25s %d\n", names(table(labels$cluster)[i]), table(labels$cluster)[i]))
}
cat(sprintf("\n"))
    
if (SEURAT_VERBOSE) {
    
  # plot variable features with and without labels
  p1 <- VariableFeaturePlot(seurat_object)
  p2 <- LabelPoints(plot   = p1, 
                    points = top_genes[1:15], 
                    repel  = TRUE)
  grid.arrange(p1, p2, nrow = 1)
  rm(p1, p2)
  
  cluster_resolution = 0.5
  
  #-------------------------------------
  source("Code/Train/seurat_analysis.R")
  seurat_object <- run_seurat_analysis(seurat_object)
  #-------------------------------------
  
  FeatureScatter(seurat_object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + NoLegend()
  VlnPlot(seurat_object, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)
  DimPlot(seurat_object, reduction = 'pca')
  DimPlot(seurat_object, reduction = 'tsne')
  DimPlot(seurat_object, reduction = 'umap', label = TRUE) + NoLegend()
  VizDimLoadings(seurat_object, dims = 1:4, reduction = 'pca')
  DimHeatmap(seurat_object, dims = 1:15, cells = 500, balanced = TRUE)
  ElbowPlot(seurat_object)
}

# Report timer
toc(log = TRUE, quiet = TRUE)
logger.info("Info: Ended classinfo%s", tic.log()[[1]])
tic.clearlog()

```


```{r}

recover = "NONE"

file_name <- file.path(rdata_path, "amb_image.rData")

if (recover == "SAVE") {
  save.image(file = file_name)
} else if (recover == "LOAD"){
  load.image(file = file_name)
}

```


There are three classes held in columns Class, Subclass and cluster
You have to choose which one to use  and copy that into the 'ident' column on position 1


# Run GBM on Class level

```{r, Class Selection}

data      <- save_data 
labels    <- save_labels 

# Start timer
logger.info("Info: Started Class level")
tic(msg = "", quiet = TRUE)

classification_category <- "Class"

#---------------------------------------------------
source("Code/Train/process_gbm_model_on_amb_data.R")
#---------------------------------------------------

p1_class            <- p1
p2_class            <- p2
cm_class            <- cm
Predicted_class     <- predicted_classes
class_summary_class <- class_summary

# Save
file_name <- file.path(rdata_path, "amb_level_class.rData")
save(cm, p1, p2, Predicted, class_summary, accuracy, medianF1, file = file_name)

# Report timer
toc(log = TRUE, quiet = TRUE)
logger.info("Info: Ended Class level%s", tic.log()[[1]])
tic.clearlog()

print (medianF1)
```
0.9996505

# Run GBM on Subclass level

```{r, Subclass Selection}

data      <- save_data 
labels    <- save_labels 

# Start timer
logger.info("Info: Started Subclass level")
tic(msg = "", quiet = TRUE)

classification_category <- "Subclass"

#---------------------------------------------------
source("Code/Train/process_gbm_model_on_amb_data.R")
#---------------------------------------------------

p1_subclass            <- p1
p2_subclass            <- p2
cm_subclass            <- cm
Predicted_subclass     <- predicted_classes
class_summary_subclass <- class_summary

# Save
file_name <- file.path(rdata_path, "amb_level_subclass.rData")
save(cm, p1, p2, Predicted_subclass, class_summary_subclass, accuracy, medianF1, file = file_name)

# Report timer
toc(log = TRUE, quiet = TRUE)
logger.info("Info: Ended Subclass level%s", tic.log()[[1]])
tic.clearlog()
print (medianF1)
```
 0.9856432

# Run GBM on cluster level

```{r, cluster Selection}

data      <- save_data 
labels    <- save_labels 

# Start timer
logger.info("Info: Started cluster level")
tic(msg = "", quiet = TRUE)

classification_category <- "cluster"

#---------------------------------------------------
source("Code/Train/process_gbm_model_on_amb_data.R")
#---------------------------------------------------

p1_cluster            <- p1
p2_cluster            <- p2
cm_cluster            <- cm
Predicted_cluster     <- Predicted
class_summary_cluster <- class_summary

# Save
file_name <- file.path(rdata_path, "amb_level_cluster.rData")
save(cm, p1, p2, Predicted, class_summary, file = file_name)

# Report timer
toc(log = TRUE, quiet = TRUE)
logger.info("Info: Ended cluster level%s", tic.log()[[1]])
tic.clearlog()

```

